\documentclass[
	%parspace, % Add vertical space between paragraphs
	%noindent, % No indentation of first lines in each paragraph
	%nohyp, % No hyphenation of words
	%twoside, % Double sided format
	%draft, % Quicker draft compilation without rendering images
	%final, % Set final to hide todos
]{elteikthesis}[2024/04/26]

% The minted package is also supported for source highlighting
% See elteikthesis_minted.tex for example
%\usepackage[newfloat]{minted}

% Document's metadata
\title{Myocardial Perfusion Imaging using Vision Transformers} % title
\date{2025} % year of defense

% Author's metadata
\author{Haris Ali}
\degree{Computer Science MSc}

% Superivsor(s)' metadata
\supervisor{Szűcs, Ádám István} % internal supervisor's name
\affiliation{PhD Candidate} % internal supervisor's affiliation
%\extsupervisor{Jane Doe} % external supervisor's name
%\extaffiliation{Senior Developer} % external supervisor's affiliation

% University's metadata
\university{Eötvös Loránd University} % university's name
\faculty{Faculty of Informatics} % faculty's name
\department{Dept. of Artificial Intelligence.} % department's name
\city{Budapest} % city
\logo{elte_cimer_szines} % logo

% Add bibliography file
\addbibresource{elteikthesis.bib}

\usepackage{cleveref}

\usepackage{tocloft}

\newlistof{equations}{equ}{List of Equations}
\newcommand{\listequationsname}{List of Equations}


\usetikzlibrary{positioning, chains, shapes.geometric, fit, shapes, arrows.meta, calc, backgrounds, intersections,shapes.arrows}

\newsavebox{\bigimage}

% The document
\begin{document}

% Set document language
%\documentlang{hungarian}
\documentlang{english}

% List of todos (not in the final document)
%\listoftodos[\todolabel]

% Title page (mandatory)
\maketitle
% Topic declaration page (mandatory) - can also be attached instead
%\includepdf{topicdeclaration.pdf}

\chapter*{Declaration}
\addcontentsline{toc}{chapter}{Declaration}

I hereby declare that this thesis titled “Myocardial Perfusion Imaging with Vision Transformers” and the work presented in it is my own original research. I confirm that:

\begin{itemize}
    \item This work was done wholly or mainly while in candidature for a research degree at Eötvös Loránd University.
    \item Where I have consulted the published work of others, this is always clearly attributed.
    \item Where I have quoted from the work of others, the source is always given.
    \item I have acknowledged all main sources of help.
    \item This thesis has not been submitted for any other degree or professional qualification.
\end{itemize}

\vspace{1cm}
\noindent
Signed: Haris Ali \\
\vspace{0.2cm}
Date: 13th April, 2025


\cleardoublepage

% Table of contents (mandatory)
\tableofcontents
\cleardoublepage

% Main content
% Abstract (mandatory)
\input{samples_en/abs.tex}
\cleardoublepage

\input{samples_en/intro.tex}
\cleardoublepage

\input{samples_en/related.tex}
\cleardoublepage

\input{samples_en/user.tex}
\cleardoublepage

\input{samples_en/results.tex}
\cleardoublepage

\input{samples_en/conc.tex}
\cleardoublepage

% Acknowledgements (optional) - in case your thesis received funding or would like to express special thanks to someone
\chapter*{\acklabel}
\addcontentsline{toc}{chapter}{\acklabel}
The research was supported by the project No. 2019-1.3.1-KK-2019-00011 financed by the National Research, Development and Innovation Fund of Hungary under the Establishment of Competence Centers, Development of Research Infrastructure Programme funding scheme. I would also like to thank my supervisor, and the open-source community for the successful completion of this project.
% Bibliography (mandatory)
\phantomsection
\addcontentsline{toc}{chapter}{\biblabel}
\printbibliography[title=\biblabel]
\cleardoublepage

% List of figures (optional) - useful over 3-5 figures
\phantomsection
\addcontentsline{toc}{chapter}{\lstfigurelabel}
\listoffigures
\cleardoublepage

% List of tables (optional) - useful over 3-5 tables
\phantomsection
\addcontentsline{toc}{chapter}{\lsttablelabel}
\listoftables
\cleardoublepage

\phantomsection
\addcontentsline{toc}{chapter}{List of Equations}
\listofequations
\cleardoublepage


\end{document}
